Neural Models for Target-Based Computer-Assisted Musical Orchestration: A Preliminary Study

### 2021-04-08 csmc_mume_2020/paper/JCMS_2021/paper.tex
Final edits before submission

After training these models on varying combinations of 10 instruments, we use perform orchestration by feeding target sounds to the networks and synthesizing a solution from the most likely samples output from the system. We then compare our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means. For quantitative comparison, we compute a specific distance metric on the amplitude spectrums of the target and solution, comparing the distance between Orchidea's solution to the target and our model's solution to the target.

@@

A complete orchestration solution, however, would normally be made by triples of instrument-pitch-dynamics. It is difficult to frame this problem as classification, since we would need to have a very high number of classes. Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in section \ref{sec:dataset}), each class would be represented by a single sample. For these reasons, our models have been trained on simplified versions of the original problem and we used different strategies to provide the remaining information needed to generate orchestrations.

Two neural classifiers have been trained on the problem of recognizing instrument-pitch pairs, which has a total of 424 classes (i.e. 424 unique instrument-pitch pairs). We want to point out, however, that this is still a simplified problem compared to assisted orchestration. Even in the case of instrument-pitch pair classification, indeed, we would miss the dynamics of the note. To cope with this problem, in the orchestration phase, we quantize the probability functions generated by the models. When a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound. Since the classes only encode instrument-pitch pairs and since we also need the dynamics, we quantize the probability function generated by the network into three echelons coresponding, respectively, to the dynamic levels \emph{pianissimo}, \emph{mezzoforte} and \emph{fortissimo} (see section \ref{sec:orchestration}). By taking the instruments and pitches that have the highest probability and by using quantized probabilities to retrieve dynamics, a full orchestrated solution can be created.  Figure \ref{fig:method} illustrates the described ideas.

%A complete orchestration solution, however, would normally be made by triples of \textit{instrument-pitch-dynamics}. It is difficult to frame this problem as classification, since we would need to have a very high number of classes. Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample. For this reasons, our models were not trained to determine \textit{instrument-pitch-dynamics} triples but instead \textit{instrument-pitch} pairs.

@@

We then computed the Mel spectrogram of the combination to be the features input to our models. We generated the Mel spectrogram using an FFT hop length of 2048 samples (the window of each FFT was $46$ms long), and a number of Mel bins equal to 128. Therefore, the Mel features fed to the model were matrices of size $128\times 345$. We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}. The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture temporal changes in the sound, assuming that musically relevant changes do not occur in a time frame less than $10$ms.

@@

\begin{table}
\begin{center}
	\begin{tabular}{|c|c|c|}
	\hline
	\textbf{Instrument} & \textbf{Abbreviation} & \textbf{Range} \\
	\hline
	Violin & Vn & G3-E7\\
	Cello & Vc & C2-C6\\
	Viola & Va & C3-C7\\
	Trumpet in C & TpC & F\#3-C6\\
	Trombone & Tbn & A\#0;A\#1-C5\\
	Oboe & Ob & A\#3-A6\\
	Horn & Hn & G1-F5\\
	Flute & Fl & B3-D7\\
	Clarinet in B$\flat$ & ClBb & D3-G6\\
	Contrabass & Cb & E1-C5\\
	Bass Tuba & BTb & F\#1-F4\\
	Bassoon & Bn & A\#1-D\#5\\
	\hline
	\end{tabular}
\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
\end{center}
\end{table}

@@

\begin{table}
  \centering
  \caption{Comparison of accuracies between SVM and RF. Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment. For the samples drawn from one of the two instruments, the pitch class of that sample is identified. For a sample not from one of the two instruments, the classifier simply attempted to identify that a sample from one of the non-specified instruments was present. In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present. \label{tab:baselines}}
    \begin{tabular}{|c|c|c|c|}
    	  \hline
      \textbf{Instr. 1} & \textbf{Instr. 2} & \textbf{SVM Acc.} & \textbf{RF Acc.}\\
      \hline
      Violin & Flute & 38.8\% & 9.8\% \\
      \hline
      Violin & Trumpet & 33.8\% & 9.1\% \\
      \hline
      Violin & Cello & 34.8\% & 6.3\% \\
      \hline
      Cello & Viola & 32.1\% & 5.8\% \\
      \hline
      Oboe & French Horn & \textbf{39.9\%} & \textbf{17.5\%} \\
      \hline
    \end{tabular}
\end{table}

%In the rest of this section, we will show the classification results of the two neural models.

@@

\begin{figure}
  \center{
  \includegraphics[width=0.8\textwidth]{../ISMIR_2020/figs/CNN_vs_ResNet.png}}
  \caption{Best overall accuracy for CNN with LSTM (50 epochs, 200k samples per epoch) and ResNet (20 epochs, 400k samples per epoch) depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
\end{figure}

@@

We calculated the distance between the target samples and our orchestrated solutions. We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions. A comparison of these results is in Table \ref{orch_eval}. While our model is not able to outperform Orchidea, it shows consistent results. We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets. Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target. You can listen to the targets and orchestrated solutions from Orchidea, ResNet, and the CNN with LSTM at \url{https://dzluke.github.io/DeepOrchestration/}.

>>>>

After training these models on varying combinations of 10 instruments, we perform orchestration by feeding target sounds to the networks and synthesizing a solution from the samples output from the system. We then compare our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means. For quantitative comparison, we compute a specific distance metric on the amplitude spectrums of the target and solution, comparing the distance between Orchidea's solution to the target and our model's solution to the target.

@@

Two neural classifiers have been trained on the problem of recognizing instrument-pitch pairs, which has a total of 424 classes (i.e. 424 unique instrument-pitch pairs). We want to point out, however, that this is still a simplified problem compared to assisted orchestration. A complete orchestration solution would normally be made by triples of instrument-pitch-dynamics. It is difficult to frame this problem as classification, since we would need to have a very high number of classes. Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in section \ref{sec:dataset}), each class would be represented by a single sample. For these reasons, our models have been trained on simplified versions of the original problem, and we employed a different strategy to determine dynamics. When a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound. Since the classes only encode instrument-pitch pairs but we also need the dynamics, we quantize the probability function generated by the network into three echelons coresponding, respectively, to the dynamic levels \emph{pianissimo}, \emph{mezzoforte} and \emph{fortissimo} (see section \ref{sec:orchestration}). By taking the instruments and pitches that have the highest probability and by using quantized probabilities to retrieve dynamics, a full orchestrated solution can be created.  Figure \ref{fig:method} illustrates the described ideas.

@@

We then computed the Mel spectrogram of the combination to be the features input to our models. We generated the Mel spectrogram using an FFT hop length of 2048 samples (the window of each FFT was $46$ms long), and a number of Mel bins equal to 128. Therefore, the Mel features fed to the model were matrices of size $128\times 345$. We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}. The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture temporal changes in the sound.

@@

\begin{table}
\begin{center}
	\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
	\begin{tabular}{|c|c|c|}
	\hline
	\textbf{Instrument} & \textbf{Abbreviation} & \textbf{Range} \\
	\hline
	Violin & Vn & G3-E7\\
	Cello & Vc & C2-C6\\
	Viola & Va & C3-C7\\
	Trumpet in C & TpC & F\#3-C6\\
	Trombone & Tbn & A\#0;A\#1-C5\\
	Oboe & Ob & A\#3-A6\\
	Horn & Hn & G1-F5\\
	Flute & Fl & B3-D7\\
	Clarinet in B$\flat$ & ClBb & D3-G6\\
	Contrabass & Cb & E1-C5\\
	Bass Tuba & BTb & F\#1-F4\\
	Bassoon & Bn & A\#1-D\#5\\
	\hline
	\end{tabular}
\end{center}
\end{table}

@@

\begin{table}
  \centering
  \caption{Comparison of accuracies between SVM and RF. Each input to the models is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.}\label{tab:baselines}
    \begin{tabular}{|c|c|c|c|}
    	  \hline
      \textbf{Instr. 1} & \textbf{Instr. 2} & \textbf{SVM Acc.} & \textbf{RF Acc.}\\
      \hline
      Violin & Flute & 38.8\% & 9.8\% \\
      \hline
      Violin & Trumpet & 33.8\% & 9.1\% \\
      \hline
      Violin & Cello & 34.8\% & 6.3\% \\
      \hline
      Cello & Viola & 32.1\% & 5.8\% \\
      \hline
      Oboe & French Horn & \textbf{39.9\%} & \textbf{17.5\%} \\
      \hline
    \end{tabular}
\end{table}

@@

\begin{figure}
  \center{
  \includegraphics[width=0.7\textwidth]{../ISMIR_2020/figs/CNN_vs_ResNet.png}}
  \caption{Best overall accuracy for CNN with LSTM (50 epochs, 200k samples per epoch) and ResNet (20 epochs, 400k samples per epoch) depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
\end{figure}

@@

We calculated the distance between the target samples and our orchestrated solutions. We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions. A comparison of these results is in Table \ref{orch_eval}. While our model is not able to outperform Orchidea, it shows consistent results. We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets. Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.
